\documentclass[10pt,a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{hyperref}

\usepackage{multicol}
\usepackage{fancyhdr}
\usepackage{enumitem}
\usepackage{tikz}
\usepackage{tikz-cd}
\usetikzlibrary{calc}
\usetikzlibrary{shapes.geometric}
\usepackage[margin=0.5in]{geometry}
\usepackage{xcolor}
\DeclareMathOperator{\RANGE}{range}
\DeclareMathOperator{\NULL}{null}

\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
    pdftitle={Tensors},
    pdfpagemode=FullScreen,
    }

%\urlstyle{same}

\newcommand{\CLASSNAME}{Functional Analysis}
\newcommand{\STUDENTNAME}{Paul Carmody}
\newcommand{\ASSIGNMENT}{Assignment \#6}
\newcommand{\DUEDATE}{May 2, 2024}
\newcommand{\SEMESTER}{Spring 2024}
\newcommand{\SCHEDULE}{T/Th 9:30 -- 10:45}
\newcommand{\ROOM}{Remote}

\pagestyle{fancy}
\fancyhf{}
\chead{ \fancyplain{}{\CLASSNAME} }
%\chead{ \fancyplain{}{\STUDENTNAME} }
\rhead{\thepage}
\input{../MyMacros}

\newcommand{\RED}[1]{\textcolor{red}{#1}}
\newcommand{\BLUE}[1]{\textcolor{blue}{#1}}
\newcommand{\GREEN}[1]{\textcolor{black!30!green}{#1}}
\newcommand{\ORANGE}[1]{\textcolor{orange}{#1}}
\newcommand{\F}{\textbf{F}}
\newcommand{\NLL}{\mathcal{N}}

\title{Advanced Linear Algebra}
\author{The Unforgetable Someone}
\date{Summer 2023}

\newcommand{\NORM}[1]{\,\left \Vert #1 \right \Vert}
%\newcommand{\HAT}[1]{}
\begin{document}

\begin{center}
	\Large{\CLASSNAME -- \SEMESTER} \\
\end{center}
\begin{center}
	\STUDENTNAME \\
	\ASSIGNMENT -- \DUEDATE\\
\end{center} 


p. 224 \#4, 7, 8, 9, 

\begin{enumerate}
	\setcounter{enumi}{3}
	\item Let $p$ be defined on a vector space $X$ and satisfy (1) and (2). Show that for any given $x_0 \in X$ there is a linear functional $\tilde{f}$ on $X$ such that $\tilde{f}(x_0)=p(x_0)$ and $|\tilde{f}(x)| \le p(x)$ for all $x \in X$.\\
	\\
	Let $f \in X'$ and $f(x_0)=p(x_0)$.  Clearly, $f$ is defined on the subspace spanned by $x_0$, that is, $f$ is linear and $f(\alpha x_0)=\alpha f(x_0)$.  The Hahn-Banach Theorem says that there exists an extension of $f$, namely, $\tilde{f} \in X'$ such that $|\tilde{f}(x)| \le p(x)$ for all $x \in X$.
	
	\setcounter{enumi}{6}
	\item Give another proof of Theorem 4.3-3 in the case of a Hilbert space.
	
	\textbf{Theorem 4.3-3a: (Bounded linear functionals, Hilbert).}  Let $X$ be a Hibert space and let $x_0\ne 0$ be any element in $X$.  Then there exists a bounded linear functional $\tilde{f}$ on $X$ such that 
	\begin{align*}
	\NORM{\tilde{f}} &= 1, &\tilde{f}(x_0)= \NORM{x_0}
	\end{align*}
	
	Proof: Let $x_0 \in X$, then $Z$ is the subspace spanned by $x_0$. Any Cauchy sequence in $Z$ will converge because that same sequence is in $X$ which is complete.  Hence. $Z$ is also complete.  We know that, for any $f_g \in Z'$ there exists a $g \in Z$ such that $f_g(x) = \ABRACKET{x,g}$ and $\NORM{f_g} = 1$.  By Hahn-Bannach, there exists an extension $\tilde{f} \in X'$ such that $\NORM{\tilde{f}} = 1$ and $|f(x_0)| = \NORM{\tilde{f}} \NORM{x_0} = \NORM{x}$.
	
	\item Let $X$ be a normed space and $X'$ its dual space. If $X \ne \{0\}$, show that $X'$ cannot be $\{0\}$.
	
	Let $f(x) = \NORM{x}$, this is linear by definition. Therefore, $f \in X'$.  We can see that when $x \ne 0$ that $f(x) \ne 0$. Therefore $f$ is not the zero function and $X \ne \{ 0 \}$.
	
	\item Show that for a separable normed space $X$, theorem 4.3-2 can be proved directly, without the use of Zorn's Lemma (which was used indirectly, namely, in the proof of Theorem 4.2-1).
	
\end{enumerate}
\newpage
p. 255 \#10, 11, 13, 14,

\begin{enumerate}
	\setcounter{enumi}{9}
	\item \textbf{(Space $c-0$)}  Let $y=(\eta_j), \eta_j \in \C$, be such that $\sum  \xi_j\eta_j$ converges for every $x=(\xi_j)\in c_0$, where $c_0\in l^\infty$ is the subspace of all complex sequences converge to zero. Show that $\sum |\eta_j|< \infty$.  (Use 4.7-3)
	
	\item Let $X$ be a Banach space, $Y$ a normed space and $T_n\in B(X,YT)$ such that $(T_nx)$ is Cauchy in $Y$ for every $x \in X$.  Show that $(\NORM{T_n})$ is bounded.
	
	\setcounter{enumi}{12}
	\item  If $(x_n)$ in a Banach space $X$ is such that $(f(x_n))$ is bounded for all $ f\in X'$, show that $(\NORM{x_n})$ is bounded.
	\item if $X$ and $Y$ are Banach spaces and $T_n\in B(X,Y), n=1,2,\cdots,$ show that equivalent statements are:
	\begin{enumerate}
		\item $(\NORM{T_n})$ is bounded.
		\item $(\NORM{T_nx})$ is bounded for all $x \in X$.
		\item $(|g(T_nx)|$ is bounded for all $x\in X$ and all $g \in Y'$.
	\end{enumerate}
\end{enumerate}
\end{document}