	\documentclass[10pt,a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{hyperref}

\usepackage{multicol}
\usepackage{fancyhdr}
\usepackage{enumitem}
\usepackage{tikz}
\usepackage{tikz-cd}
\usetikzlibrary{calc}
\usetikzlibrary{shapes.geometric}
\usepackage[margin=0.5in]{geometry}
\usepackage{xcolor}
\DeclareMathOperator{\RANGE}{range}
\DeclareMathOperator{\NULL}{null}

\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
    pdftitle={Tensors},
    pdfpagemode=FullScreen,
    }

%\urlstyle{same}

\newcommand{\CLASSNAME}{Functional Analysis}
\newcommand{\STUDENTNAME}{Paul Carmody}
\newcommand{\ASSIGNMENT}{Assignment \#5}
\newcommand{\DUEDATE}{April 18, 2024}
\newcommand{\SEMESTER}{Spring 2024}
\newcommand{\SCHEDULE}{T/Th 9:30 -- 10:45}
\newcommand{\ROOM}{Remote}

\pagestyle{fancy}
\fancyhf{}
\chead{ \fancyplain{}{\CLASSNAME} }
%\chead{ \fancyplain{}{\STUDENTNAME} }
\rhead{\thepage}
\input{../MyMacros}

\newcommand{\RED}[1]{\textcolor{red}{#1}}
\newcommand{\BLUE}[1]{\textcolor{blue}{#1}}
\newcommand{\GREEN}[1]{\textcolor{black!30!green}{#1}}
\newcommand{\ORANGE}[1]{\textcolor{orange}{#1}}
\newcommand{\F}{\textbf{F}}

\title{Advanced Linear Algebra}
\author{The Unforgetable Someone}
\date{Summer 2023}

\newcommand{\NORM}[1]{\,\left \Vert #1 \right \Vert}
\begin{document}

\begin{center}
	\Large{\CLASSNAME -- \SEMESTER} \\
\end{center}
\begin{center}
	\STUDENTNAME \\
	\ASSIGNMENT -- \DUEDATE\\
\end{center} 

p. 200 \#2,3,4,5,6,10. 
\begin{enumerate}
	\setcounter{enumi}{1}
	\item Let $H$ be a Hilbert space and $T: H \to H$ a bijective bounded linear operator whose inverse is bounded.  Show that $(T^*)^{-1}$ exists and 
	\begin{align*}
		(T^*)^{-1}=(T^{-1})^*
	\end{align*}
	
	\begin{align*}
		(TT^{-1})^* &= (T^{-1})^*T^*\\
		TT^{-1} &= T^*(T^*)^{-1} \\
		(T^{-1})^*T^* &= ((T^*)^{-1})^*T^{**}
	\end{align*}
	
	\item if $(T_n)$ is a sequence of bounded linear operators on a Hilbert space adn $T_n \to T$, show that $T_n^*\to T^*$.
	
	\item Let $H_1$ and $H_2$ be Hilbert spaces and $T: H_1 \to H_2$ a bounded linear operator.  If $M_1\subset H_1$ and $M_2 \subset H_2$ are such that $T(M_1) \subset M_2$, show that $M_1^\perp \subset T^*(M_2^\perp)$.
	
	\item Let $M_1$ and $M_2$ in Prov. 4 be closed subspaces.   Show that then $T(M_1)\subset M_2$ if and only if $M_1^\perp \supset T^*(M_2^\perp)$.
	
	\item If $M_1=\mathcal{N}(T)=\{x\,|\,Tx=0\}$ in Prob. 4, show that 
	\begin{enumerate}
		\item $T^*(H_2) \subset M_1^\perp$
		\item $[T(H_1)]^\perp \subset \mathcal{N}(T^*)$
		\item $M_1=[T^*(H_2)]^*$
	\end{enumerate}
	\setcounter{enumi}{9}
	\item \textbf{(Right shift operator)} Let $(e_n)$ be a total orthonormal sequence in a separable Hilbert space $H$ and define the \textit{right shift operator} to be the linear operator $T: H \to H$ such that $Te_n=e_{n+1}$ for $n=1,2,\cdots$.  Explain the name.  Find the range, null space, norm and Hilbert-adjoint operator of $T$.

\end{enumerate}

\newpage
p. 207 \#4, 5

\begin{enumerate}
	\setcounter{enumi}{3}
	\item Show that for any bounded linear operator $T$ on $H$, the operators 
	\begin{align*}
		T_1=\frac{1}{2}(T+T^*) \text{  and  } T_3=\frac{1}{2i}(T-T^*)
	\end{align*}are self-adjoint.  Show that 
	\begin{align*}
		T=T_1+iT_2 \text{      } T^*=T_1=iT_2.
	\end{align*}Show uniquness, that is, $T_1+iT_2=S_1+iS_2$ implies $S_i=T_i$ and $S_2=T_2$; here, $S_1$ and $S_2$ are self-adjoint by assumptioin.
	
	\item On $\C^2$ (cf. 3.1-4) let the operator $T: \C^2\to \C^2$ be defined by $Tx=(\xi_1+i\xi_2, \xi-i\xi_2),$ where $x=(\xi_1,\xi_2)$.  Find $T^*$.  Show that we have $T^*T=TT^*=2I$.  Find $T_1$ and $T_2$ as defined in prob. 4.

\end{enumerate}
\end{document}