\documentclass[10pt,a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{hyperref}

\usepackage{multicol}
\usepackage{fancyhdr}
\usepackage[inline]{enumitem}
\usepackage{tikz}
\usepackage{tikz-cd}
\usetikzlibrary{calc}
\usetikzlibrary{shapes.geometric}
\usepackage[margin=0.5in]{geometry}
\usepackage{xcolor}

\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
    pdftitle={Tensors},
    pdfpagemode=FullScreen,
    }

%\urlstyle{same}

\newcommand{\CLASSNAME}{Math 5110 -- Real Analysis I}
\newcommand{\STUDENTNAME}{Paul Carmody}
\newcommand{\ASSIGNMENT}{Homework \#1 }
\newcommand{\DUEDATE}{September 2, 2024}
\newcommand{\SEMESTER}{Fall 2024}
\newcommand{\SCHEDULE}{MW 11:00 -- 12:15}
\newcommand{\ROOM}{Remote}

\newcommand{\MMN}{M_{m\times n}}
\newcommand{\FF}{\mathcal{F}}

\pagestyle{fancy}
\fancyhf{}
\chead{ \fancyplain{}{\CLASSNAME} }
%\chead{ \fancyplain{}{\STUDENTNAME} }
\rhead{\thepage}
\input{../MyMacros}

\newcommand{\RED}[1]{\textcolor{red}{#1}}
\newcommand{\BLUE}[1]{\textcolor{blue}{#1}}

\begin{document}

\begin{center}
	\Large{\CLASSNAME -- \SEMESTER} \\
	\large{ w/Professor Liu}
\end{center}
\begin{center}
	\STUDENTNAME \\
	\ASSIGNMENT -- \DUEDATE\\
\end{center} 

\begin{enumerate}[label=\Roman*.]
\item This problem review continuity for functions on real line.

We say a function $f: \R \to \R$ is \textit{continuous} at a point $a \in \R$ if for any $\epsilon > 0$, there is a $\delta > 0$ such that $|x-a| < \delta$ implies $|f(x)-f(a)|< \epsilon$.
\begin{enumerate}[label=(\alph*)]
\item Show that $f(x)=x^2$ is continuous at $x=2$.

\RED{Given an $\epsilon > 0$, when $|f(x)-4|< \epsilon$, $ |x^2 -4| < \epsilon$.  Let $\delta < \sqrt{\epsilon + 4}$ \\
If \begin{align*}
	(2+\delta)^2-4 &< \epsilon \\
	(2+\delta)^2&< \epsilon +4\\
	2+\delta < \sqrt{\epsilon +4} \\
	\delta < \sqrt{\epsilon+4}-2
\end{align*} }

\item Suppose that $f$ is continuous at $a$ and $f(a) \ne 0$. Show that $f$ is nonzero in some open interval containing $a$.

\BLUE{Since $f$ is continuous at $a$ and $f(a) \ne 0$ then for every $\epsilon > 0$ such that when $|f(x)-f(a)| < \epsilon$ or $f(x) = f(a) \pm \epsilon$. Without loss of generality, we can also say $f(a) \ne \epsilon$. Since both $f(a)$ and $\epsilon$ are both non-zero, $f(x) \ne 0$. }

\end{enumerate}

\item This problem review derivatives.
\begin{enumerate}[label=(\alph*)]
\item Let $f(x)=x^n$ for some positive integer $n$.  Using the definition of the derivative, and the binomial theorem, show that $f'^{n-1}$.
\item Is the function 
\begin{align*}
	f(x) &= \BINDEF{x^2, & x \le 0,}{-x^2, & x \le 0}
\end{align*}differentiable at $x=0$.

\BLUE{
\begin{align*}
	\lim_{x \to 0^-} f(x) &= \lim_{x \to 0^-} x^2 = 0 \\
	\lim_{x \to 0^+} f(x) &= \lim_{x \to 0^+} x^2 = 0 \\
	\lim_{x \to 0^-} f(x) &= \lim_{x \to 0^+} f(x)
\end{align*}$f(x)$ is both continuous and differentiaable at $x=0$. }
\end{enumerate}

\newpage
\item This problem reviews $\sup$ and $\inf$.

For any subset $A \subset \R$, we say that $M$ is an \textit{upper bound} for $A$ if $x \le M$ for all $x \in A$.  If a set $A$ has a finite upper bound, we say it is \textit{boundared above}.  It is a theorem about the set $\R$ that \textit{for any set $A \subset \R$ that is bounded above, there exists a least (smallest) upper bound for $A$.}  This least upper bound is called supermum of $A$, and denoted $\sup A$.  By definition, the number $\sup A$ has two properties.
\begin{enumerate}[label = (\roman*)]
\item $x \le \sup A$ for all $x \in A$ (i.e., $\sup A$ is an upper bound for $M$).
\item for any $M$ that is an upper bound for $A$, we have $\sup A \le M$.
\end{enumerate}For sets that are not bounded above, we say that $\sup A = + \infty$.  we often write things like 
\begin{align*}
	\sup_{x \in A} f(x),
\end{align*}to denote the supremum of the set $\{ f(x): x \in A\}$, where $f$ is a some function.

Similarly, any set that is bounded below has a \textit{greatest lower bound} called the \textit{infimum}, denoted $\inf A$.  It satisfies the same properties as $\sup A$ with the inequalities reversed.
\begin{enumerate}[label=(\alph*)]

\item Find $\sup A$ and $\inf A$ for $A = (1,2]$, $A=\{1,\frac{1}{2},\frac{1}{3}, \cdots \}$, and $A = \{0,1,2,3,\dots\}$.

\begin{itemize}
\item $A = (1,2]$, \BLUE{ $\sup A = 2$ and $\inf A = 1$ }
\item $A=\{1,\frac{1}{2},\frac{1}{3}, \cdots \}$, \BLUE{ $\sup A = 1$, and $\inf A = \LIM{n\to \infty} \frac{1}{n} = 0$.}
\item $A = \{0,1,2,3,\dots\}$.  \BLUE{ $\sup A = \LIM{n\to \infty} = \infty$, and $\inf A = 0$}
\end{itemize}

\item Find $\SUP{x \in (0, 1)}(1+x^2)^{-1}$

\BLUE{ Let $f(x) = (1+x^2)^{-1}$.  On the interval $(0,1)$ we can see that it is strictly decreasing, that is $a<b \implies f(a)> f(b)$.  Thus, $\SUP{x \in (0, 1)}f(x) = f(0) = (1+0^2)^{-1} = 1$.}

\item Assume that $\sup A < \infty$, and show that for every $\epsilon > 0$, there exists $x \in A$ such that $x > \sup A - \epsilon$.

\BLUE{Given any $\epsilon > 0$ let $x > \sup A - \epsilon$.  If $x \not \in A$ then $x$ is an upper bound of $A$, i.e., $x \in M$ and $x< \sup A$, but that violates proper (ii).  Hence, $x \in A$.} 

\item For any two functions $f,g : \R \to \R$, and any set $A \subset \R$, show that $\SUP{x \in A}(f(x)+g(x)) \le \SUP{x\in A} f(x) + \SUP{x\in A} g(x)$.

\BLUE{
\begin{align*}
	f(x) &\le \SUP{x\in A} f(x) \AND g(x) \le \SUP{x\in A} g(x), \forall x \in {x \in A}\\
	\therefore f(x) + g(x) &\le \SUP{x\in A} f(x) + \SUP{x\in A} g(x), \forall x \in A \\
	\AND \SUP{x \in A}(f(x)+g(x)) &\le \SUP{x\in A} f(x) + \SUP{x\in A} g(x)
\end{align*} }

\end{enumerate}

\newpage
\item Section 1.1, Exercise 5, 6, 13.

\begin{description}
\item \textit{Exercise 1.1.5}. Let $n \ge 1$, and let $a_1, a_2, \dots, a_n$ and $b_1, b_2, \dots, b_n$ be real numbers verify the identity
\begin{align*}
	&\PAREN{\sum_{i=1}^na_ib_i}^2+\frac{1}{2}\sum_{i=1}^n\sum_{j=1}^n(a_ib_j-a_jb_i)^2= \PAREN{\sum_{i=1}^na_i^2}\PAREN{\sum_{j=1}^n b_j^2}, &(1.3)
\end{align*}and conclude \textit{Cauchy-Schwarz inequality}
\begin{align*}
	\BARS{\sum_{i=1}^n a_1b_i} \le \PAREN{\sum_{i=1}^na_i^2}^{1/2}\PAREN{\sum_{j=1}^nb_j^2}^{1/2}
\end{align*}Then use the Cauchy-Schwarz inequality to prove the \textit{triangle inequality}
\begin{align*}
	\PAREN{\sum_{i=1}^n(a_i+b_i)^2}^{1/2} \le \PAREN{\sum_{i=1}^n a_i^2}^{1/2}+\PAREN{\sum_{i=j}^n b_j^2}^{1/2}
\end{align*}

\RED{Let's start by expanding the center term\begin{align*}
	\sum_{i=1}^n\sum_{j=1}^n(a_ib_j-a_jb_i)^2 &= \sum_{i=1}^n\sum_{j=1}^n ( (a_ib_j)^2+(a_jb_i)^2 -(a_ib_ja_jb_i)^2)\\
	&= \sum_{i=1}^n a_i^2\sum_{j=1}^n b_j^2 + \sum_{i=1}^nb_i^2\sum_{j=1}^n a_i^2 - 2\sum_{i=1}^na_ib_i\sum_{j=1}^na_jb_j \\
	&= 2\PAREN{\sum_{i=1}^na_i^2}\PAREN{\sum_{j=1}^nb_j^2}-2\PAREN{\sum_{i=1}^na_ib_i}^2
\end{align*}Equation 1.3 then becomes
\begin{align*}
	\PAREN{\sum_{i=1}^na_ib_i}^2+\frac{1}{2}\sum_{i=1}^n\sum_{j=1}^n(a_ib_j-a_jb_i)^2= \PAREN{\sum_{i=1}^na_i^2}\PAREN{\sum_{j=1}^n b_j^2} \\
	\PAREN{\sum_{i=1}^na_ib_i}^2+\PAREN{\sum_{i=1}^na_i^2}\PAREN{\sum_{j=1}^nb_j^2}-\PAREN{\sum_{i=1}^na_ib_i}^2= \PAREN{\sum_{i=1}^na_i^2}\PAREN{\sum_{j=1}^n b_j^2}
\end{align*}which is true.  Since
\begin{align*}
	\PAREN{\sum_{i=1}^na_ib_i}^2+\frac{1}{2}\sum_{i=1}^n\sum_{j=1}^n(a_ib_j-a_jb_i)^2&= \PAREN{\sum_{i=1}^na_i^2}\PAREN{\sum_{j=1}^n b_j^2} \\
	\PAREN{\sum_{i=1}^na_ib_i}^2&=\PAREN{\sum_{i=1}^na_i^2}\PAREN{\sum_{j=1}^n b_j^2} - \frac{1}{2}\sum_{i=1}^n\sum_{j=1}^n(a_ib_j-a_jb_i)^2 \\
	\therefore \BARS{\sum_{i=1}^na_ib_i} &\le \PAREN{\sum_{i=1}^na_i^2}^{1/2}\PAREN{\sum_{j=1}^n b_j^2}^{1/2}
\end{align*}Let's start by taking the square of the distance from $a+b$ to zero using the $\ell^2$.
\begin{align*}
	d_{\ell^2}(a+b, 0)^2 &= \sum_{i=1}^n (a_i+b_i)^2 \\
	&= \sum_{i=1}^n (a_i^2 + b_i^2 + 2a_ib_i) \\
	&= \sum_{i=1}^n a_i^2 + \sum_{i=1}^nb_i^2 + 2\sum_{i=1}^na_ib_i 
\end{align*}apply Cauchy-Schwarz and factor.
\begin{align*}
	d_{\ell^2}(a+b, 0)^2 &\le d_{\ell^2}(a,0) + d_{\ell^2}(b, 0) + 2 \PAREN{\sum_{i=1}^na_i^2}^{1/2}\PAREN{\sum_{j=1}^n b_j^2}^{1/2} \\
	&\le d_{\ell^2}(a,0) + d_{\ell^2}(b, 0) + 2\PAREN{ d_{\ell^2}(a,0)\cdot d_{\ell^2}(b,0) }^{1/2} \\
	&\le \PAREN{ d_{\ell^2}(a,0)^{1/2} + d_{\ell^2}(b,0)^{1/2}}^2
\end{align*}Expand the $\ell^2$ metrics and take the square root of both sides and 
\begin{align*}
	\PAREN{\sum_{i=1}^n(a_i+b_i)^2}^{1/2} &\le \PAREN{\sum_{i=1}^n a_i^2}^{1/2}+\PAREN{\sum_{i=j}^n b_j^2}^{1/2}
\end{align*} }

\newpage
\item \textit{Exercise 1.1.6}  Show that $(\R^n, d_{l^2})$ in Example 1.1.6 is indeed a metric space. (Hint: use Exercise 1.1.5)\\
\textbf{Example 1.1.6} (Euclidean spaces). Let $n \ge 1$ be a natural number, and let $\R^n$ be the space of $n$-tupes of real numbers:
\begin{align*}
	\R^n = \{(x_1,x_2, \dots, x_n): x_1, \dots, x_n \in \R\}
\end{align*}We define the \textit{Euclidean metric} (also called the\textit{ $l^2$ metric}) $d_{l^2}: \R^n \times \R^n \to \R$ by
\begin{align*}
	d_{l^2} ((x_1, \dots, x_n), (y_1, \dots, y_n)) &= \sqrt{(x_1-y_1)^2+\cdots+ (x_n-y_n)^2}\\
	&= \PAREN{\sum_{i=1}^n (x_i-y_i)^2 }^{1/2}
\end{align*}

\BLUE{We must prove that $d_{\ell^2}$ is symmetric, positive definite and that the triangle inequality holds.
\begin{description}
	\item Symmetric: show that $d_{\ell^2}(x,y) = d_{\ell^2}(y,x)$.
		\begin{align*}
			d_{l^2} ((x_1, \dots, x_n), (y_1, \dots, y_n)) &= \PAREN{\sum_{i=1}^n (x_i-y_i)^2 }^{1/2} \\
			&= \PAREN{\sum_{i=1}^n (y_i-x_i)^2 }^{1/2} \\
			&= d_{l^2} ((y_1, \dots, y_n), (x_1, \dots, x_n))
		\end{align*}
	\item Positive Definite: show that $d_{\ell^2}(x,y)\ge 0$ and $d_{\ell^2}(x,y) = 0 \to x=y$.\\
	The square root is taken as a positive value.  $d_{l^2} ((x_1, \dots, x_n), (y_1, \dots, y_n))=0$ implies that each $x_i - y_i=0$ therefore $x=y$.
	\item Triangle Inequality: show that $d_{\ell^2}(x,z) \le d_{\ell^2}(x,y)+d_{\ell^2}(y,z)$\\
	Excercise 1.1.5 proves the triangulate inequality replacing $a_i = x_i$ and $b_i = y_i$.
\end{description}
}

\newpage
\item \textit{Exercise 1.1.13} Prove Proposition 1.1.19.
\textbf{Proposition 1.1.19} (Convergence in a the discrete metric).  \textit{Let $X$ be any set, and let $d_{\text{disc}}$ be the discrete metric on $X$.  let $(x^{(x=n)})_{n=m}^\infty$ be a sequence of points in $X$, and let $x$ be a point in $X$. Then $(x^{(n)})_{n=m}^\infty$ convergent to $x$ with respect to the discrete metric $d_{\text{disc}}$ if and only if there exists $N \ge m$ sucht ath $x^{(n)}=x$ for all $n \ge N$. }

\item For this problem only, you do not need to give proofs.  just write the answers.

For each set, idenitfy the boundary, interior, and closure of $A$, and say whether $A$ is open, closed, both or neither.  We are working in $\R^2$ with the standard distance.  Unless othewise noted, the ambient space is $\R^2$.
\begin{enumerate}[label=(\alph*)]
\item $A = \{(x_1,x_2) \in \R^2: x_1 < 1\}$
\item $A = \{(1/n, 2/n) : n=1,2,3,\dots\}$ (Note: $(1/n, 2/n)$ is a vector in $\R^2$, not an open intevarl in $\R$.).
\item $A \ \{x=(x_1,x_2) \in \R^2: x_1 > 0, d(x,0) \le 1\}$, in the relative topology with respect to $Y = \{(x_1, x_2) \in \R^2: x_1 > 0\}$.

\end{enumerate}

\item Let $(X, d)$ be a metric space.
\begin{enumerate}[label=(\alph*)]

\item For a given piont $x_0 \in X$, show the singleton set $\{x_0\}$ is closed.
\item Let $x_0 \in X$ and $r > 0$. Show that the ball
\begin{align*}
	B(x_0, r) &= \{x \in X : d(x, x_0) < r\}
\end{align*}is open

\end{enumerate}

\end{description}

\end{enumerate}

\end{document}