\documentclass[10pt,a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{hyperref}

\usepackage{multicol}
\usepackage{fancyhdr}
\usepackage[inline]{enumitem}
\usepackage{tikz}
\usepackage{tikz-cd}
\usetikzlibrary{calc}
\usetikzlibrary{shapes.geometric}
\usepackage[margin=0.5in]{geometry}
\usepackage{xcolor}

\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
    pdftitle={Tensors},
    pdfpagemode=FullScreen,
    }

%\urlstyle{same}

\newcommand{\CLASSNAME}{Math 5110 -- Real Analysis I}
\newcommand{\STUDENTNAME}{Paul Carmody}
\newcommand{\ASSIGNMENT}{Homework \#2 }
\newcommand{\DUEDATE}{September 18, 2024}
\newcommand{\SEMESTER}{Fall 2024}
\newcommand{\SCHEDULE}{MW 11:00 -- 12:15}
\newcommand{\ROOM}{Remote}

\newcommand{\MMN}{M_{m\times n}}
\newcommand{\FF}{\mathcal{F}}

\pagestyle{fancy}
\fancyhf{}
\chead{ \fancyplain{}{\CLASSNAME} }
%\chead{ \fancyplain{}{\STUDENTNAME} }
\rhead{\thepage}
\input{../MyMacros}

\newcommand{\RED}[1]{\textcolor{red}{#1}}
\newcommand{\BLUE}[1]{\textcolor{blue}{#1}}

\begin{document}

\begin{center}
	\Large{\CLASSNAME -- \SEMESTER} \\
	\large{ w/Professor Liu}
\end{center}
\begin{center}
	\STUDENTNAME \\
	\ASSIGNMENT -- \DUEDATE\\
\end{center} 

\begin{enumerate}[label=\Roman*.]
\item Consider a sequence $x_n$ of real numbers.  The \textit{limit inferior} and \textit{limit superior} of $x_n$ are defined by 
\begin{align*}
	\LIMINFN x_n = \lim_{n->\infty}\PAREN{\inf_{k \ge n} x_k }, \,\, \LIMSUPN x_n = \lim_{n->\infty}\PAREN{\sup_{k \ge n} x_k }
\end{align*}
	\begin{enumerate}[label=(\alph*)]
		\item Show that
			\begin{align*}
				\LIMINFN x_n = \sup_{n\ge 0} \PAREN{\inf_{k\ge n} x_k}
			\end{align*}and
			\begin{align*}
				\LIMSUPN x_n = \inf_{k\ge 0}\PAREN{\sup_{k\ge n} x_k}
			\end{align*}
		
		\BLUE{First, $\LIMINFN$: Let $y_n = \underset{k\ge n}\inf\, x_k$. Then, given any $j>k, y_k \le y_j$.  That is, $y_n$ is a bounded increasing sequence.  All $y_n \le \underset{n \to \infty}\sup y_n$  Thus, the $\LIMN y_n = \underset{n\ge 0}\sup\, y_n$.\\
		Next, $\LIMSUPN$:  Let $z_n = \underset{k\ge n}\sup\, x_k$. Then, given any $j>k, z_k \ge z_j$.  That is, $z_n$ is a bounded decreasing sequence.  All $z_n \ge \underset{n \to \infty}\inf z_n$  Thus, the $\LIMN z_n = \underset{n\ge 0}\inf\, z_n$.\\
		}
		
		\item Show that $\LIMINFN x_n$ and $\LIMSUPN x_n$ are well-defined for any sequence $x_n$.  (Unlike $\LIM{n\to\infty} x_n$.)  We allow values of $\infty$ and $-\infty$

		\BLUE{Using $(y_n)$ from (a), that must exist one and only one value for $\LIMN y_n$ as it is bounded and increasing, thus its limit is well-defined.  Similarly, for $(z_n)$.
		}
				
		\item Let $x_n$ be a bounded sequence, and let $L$ be the set of limit points of $x_n$, i.e., the set of all limits of subsequences of $x_n$.  Show $\LIMINFN x_n = \inf L$ and $\LIMSUPN = \sup L$.
		
		\BLUE{Let $L$ be the set of limit points for $x_n$.  Then, for any $w \in L$ there is a $(w_k)\in (x_n)$ subsequence such that $\underset{k\to \infty}\lim w_k = w$.  The $\underset{k\to\infty}\inf w_k \ge \inf L \ge \LIMINFN x_n.$  However, from (a) we can see that 
		\begin{align*}
				\LIMINFN x_n = \sup_{n\ge 0} \PAREN{\inf_{k\ge n} x_k}		
		\end{align*}therefore $\LIMINFN x_n \ge \inf L$ thus $\LIMINFN x_n = \inf L$.\\
		Similarly, for $\LIMSUPN x_n$.
		}
		\item Let $x_n$ be a bounded sequence.  Conclude using (c) that $\LIMINFN x_n \le \LIMSUPN x_n$, with equality if and only if $x_n$ is convergent.
		
		\BLUE{By definition, $\inf L \le \sup L$ therefore $\LIMINFN x_n \le \LIMSUPN x_n$.  Therefore, from (a)
		\begin{align*}
				\sup_{n\ge 0} \PAREN{\inf_{k\ge n} x_k} \le \inf_{n\ge 0}\PAREN{\sup_{k\ge n} x_k}
		\end{align*}Now using $(y_n)$ and $(z_n)$ from (a) we can see that we have \begin{align*}
			\sup_{n\ge 0} \,y_n \le \inf_{n\ge 0}\,z_n
		\end{align*}we have a bounded increasing sequence on the left less than a bounded decreasing sequence on the right.  They can only be equal if they converge to the same value.
		}
	\end{enumerate}
	
	\item Prove that for any (possibly uncountable) collection $(F_\alpha)_{\alpha\in A}$ of closed sets, the intersection $F = \bigcup_{\alpha \in A} F_\alpha$ is closed, in two ways.
	\begin{enumerate}[label=(\alph*)]
		\item Using the fact that any union of open sets is open, and DeMorgan's Laws from set theory, which state
		\begin{align*}
			X\backslash \PAREN{\bigcup_{\alpha\in A} E_\alpha } = \bigcap_{\alpha \in A} \PAREN{X \backslash E_\alpha} \AND X\backslash\PAREN{\bigcap_{\alpha \in A} E_\alpha} = \bigcup_{\alpha\in A} \PAREN{X \backslash E_\alpha}
		\end{align*}for all collection of sets $(E_\alpha)_{\alpha\in A}$
		
		\BLUE{Given that every open set, $E\in X$ is the union of other open sets $\bigcup_{\alpha \in A} E_\alpha$ for some index set $A$ (whether countable or uncountable).  We know that the complement is closed and the complement can be expressed as
		\begin{align*}
			E^c &= X\backslash E \\
			&= X\backslash \PAREN{\bigcup_{\alpha\in A} E_\alpha }\\
			&= \bigcap_{\alpha \in A} \PAREN{X \backslash E_\alpha}
		\end{align*}each $E_\alpha$ is the complement of an open set, hence they are closed.  Thus, $E^c$ which is closed is made up of the intersection of closed sets.
		}
		
		\item More directly, using the fact that a set $G$ is closed if and only if for any convergent sequence $(x_n)$ with all $x_n \in G$, the limit $x$ is also in $G$.
		
		\BLUE{Let $F, G \in X$ be closed sets and let $(x_n) \subset G$ and $(y_n) \subset F$ both be convergent sequences. Further, we let $(x_n), (y_n) \subset G\cap F$. Not that $F$ closed means that $(x_n) \in F$ implies that $\LIMN x_n \in F$, thus $\LIMN x_n \in G \cap F$ and a similar argument can be made for $y_n$ and $G$.  Thus sequences contained in $G\cap F$ must also contain their limits and $G\cap F$ is closed.  This can extend to any number of intersections.
		}
	\end{enumerate}
	
	\item\begin{enumerate}[label=(\alph*)]
		\item Let $(x_n)$ be a Cauchy sequence in a metric space $X$.  Show that if a subsequence $(x_{n_j})$ of $x_n$ converges to $x$, then the entire sequence also converges to $x$.
		
		\BLUE{Let $(x_n)$ be Cauchy and let $(x_{n_j})$ be a convergent subsequence of $(x_n)$.  Then, there exists for every $\epsilon > 0$ there is $N \in \N$ such that whenever $j,k > N, |x_{n_j} - x_{n_k} | < \epsilon$.  Let $M = \min\{n_j, n_k\}$.  We can see that $|x_m-x_k| < \epsilon$.  $x_m, x_k \in x_n$ and $x_n$ is Cauchy, therefore all this is true for all elements of $m,k >M$, hence $(x_n)$ converges.
		}
		\item Show that the metric space
		\begin{align*}
			C^1\PAREN{(-1,1)}=\{f:(-1,1)\to\R, f \text{ is differentiable and }f'\text{ is continuous in }(1,-1)\}
		\end{align*}with the metric
		\begin{align*}
			d(f,g) = \sup_{x\in (-1,1)} |f(x)-g(x)|
		\end{align*}is not complete.  (Hint: similar to the proof that the rational numbers are not complete, find a sequence $C^1\PAREN{(-1,1)}$ that converges in $d$ metric to a function that is not in $C^1\PAREN{(-1,1)}$, and show that this sequence is Cauchy.)
		
		\BLUE{Let $f_n(x) = x^{\frac{1}{2n+1}}$. We can see that given any $\epsilon > 0$ there is $N \in \N$ such that $n,m >N$ the distance 
		\begin{align*}
			d(f_n,f_m)&=\sup_{x\in (-1,1)} |f_n(x)-f_m(x)| \\
			&=\sup_{x\in (-1,1)} |x^{1/2n+1}-x^{1/2m+1}| \\
		 	&< \epsilon
		\end{align*}The functions are all differentiable and their derivatives are continuous, but
		\begin{align*}
				\lim_{n\to\infty} x^{\frac{1}{2n+1}} = \BINDEF{-1 & x<0}{1 & x>0}, \forall x \in (-1,1)
		\end{align*}which is not a member of $C^1\PAREN{(-1,1)}$
		}
	\end{enumerate}
	
	\item Let $A$ and $B$ be subsets of the metric space $X$.  which one of the following is true?
	\begin{align}
		(A\cup B)^o &= A^o \cup B^o, &\label{eq:eq1}\tag{2.1}\\
		(A \cup B)^o &\subset A^o\cup B^o, & \text{"=" fails for some $A$ and $B$} \label{eq:eq1}\tag{2.2} \\
		(A\cup B)^o &\supset A^o\cup B^o, &\text{"=" fails for some $A$ and $B$}\label{eq:eq1}\tag{2.3}
	\end{align}
	
	\BLUE{(2.3) Consider $X = \R^3$ and $A$ is the open unit disc in the X-Y plane centered at the origin and $B$ is the open unit disc in the Y-Z plane centered at the origin.  $(A \cup B)^o \supset A^o\cup B^o$. }
	
	\item Let $C^0([a,b])$ be the space of continuous functions on $[a,b]$, with the metric $d(f,g) = \SUP{x\in [a,b]}|f(x)-g(x)|$.
	
	Show that the map $I: C^0([a,b]) \to \R$ defined by $I(f)=\int_a^b f(x) dx$ is continuous mapping from $C^0([a,b])$ to $\R$.
	
	\BLUE{$I$ is continuous if for ever $\epsilon > 0$ there exists $\delta > 0$ such that $d(I(f),I(g))<\epsilon$ whenever $d(f,g) < \delta$. Or 
	\begin{align*}
		d(I(f),I(g)) &= \SUP{x\in [a,b]}|I(f(x))-I(g(x))|\\
			&= \SUP{x\in [a,b]}\BARS{\int_a^b f(x) dx-\int_a^b g(x) dx}\\
			&= \SUP{x\in [a,b]}\BARS{\int_a^b f(x) - g(x) dx}\\
			&= \SUP{x\in [a,b]}\int_a^b \BARS{f(x) - g(x)} dx\\
			&\le \int_a^b \SUP{x\in [a,b]}\BARS{f(x) - g(x)} dx\\
			&\le \int_a^b d(f,g) dx\\
			&\le d(f,g)[b-a]
	\end{align*}Thus when $\epsilon > 0$ choose $\delta > [b-a]d(f,g)$.  Hence, $I$ is continuous.
	}
	
	\item \textbf{Proposition 2.3.2} (Maximum principle). \textit{ Let $(X,d)$ be a compact metric space, and let $f:X \to \R$ be a contnuous function.  Then $f$ is bounded.  Furthermore, $f$ attains its maximum at some point $x_{\max} \in X$, and also attains its minimum at some point $x_{\min} \in X$. }. \\
	Prove Propostion 2.3.2 in the text, in two different ways.:
	\begin{enumerate}[label=\alph*)]
		\item As a consequence of Theorem 2.3.1 in text.
		
		\BLUE{Let $f: X \to \R$ be a continuous function on a compact set $X$.  Then, by 2.3.1, $f(X)$ is a compact set.  Every compact set in $\R$ is an interval.  Let $\ABRACKET{a,b}$ be that interval, that is, $f: X \to \ABRACKET{a,b}$.  If $f$ were unbounded, then there would exist an $x \in X$ such that $f(x) \not \in \ABRACKET{a,b}$ which cannot happen.  Therefore, there must exists values in the domain $x_{\min}$ and $x_{\max}$ which are the maximum and minimum values of $f$, namely, $a, b$, respectively.
		}
		
		\item Directly, using the sequential definition of compactness.
		
		\BLUE{Let $(x_n) \in X$ be any sequence in the compact space $X$.  Being compact, $(x_n)$ must converge and $\LIMN x_n = x \in X$.  Let $f: X \to \R$ be a continuous function.  $x_n$ converges implies that $f(x_n)$ also converges.  Therefore, $\LIMN f(x_n) = f(x)$ and is finite (otherwise $f$ would not be continuous).  Therefore, there exists an upper and lower bound of $f$.  Let $L$ be the lower bound and $(y_n) \in f(X)$ be a sequence such that $\LIMN y_n = L$.  Then, let $z_i$ be such that $f(z_i) = y_i$ for all $i$.  Then, we have a sequence $(z_n)\in X$ which must converge.  Thus, $\LIMN f(z_n) = L$ and $\LIMN z_n = x_{\min}$.  \\Similarly, for $x_{\max}$.
		}

	\end{enumerate}
	
	\item Let $f:\R^n \to \R$ be a continuous function such that 
	\begin{align*}
		\lim_{|x| \to \infty} f(x) &= + \infty
	\end{align*}Prove that $f$ attains its minimum.
	
	Recall that by definition, the limit in (??) means that Given $A > 0$, there is $R>0$ such that 
	\begin{align*}
		f(x) >A \text{    for all } x \not\in B_R
	\end{align*}in other words, $f(x) >A$ whenever $|x| \ge R$. Here, $|x| = d_2(x,0)$ and $d_2$ is the standard Euclidean distance defined in Example 1.4.
	
	\BLUE{Given any $A> 0$ there exists an $R>0$ such that $f(x) >A$ whenever $|x| > R$.  Therefore, $f(x) \le A$ whenever $|x| < R$.  $f(x)$ is bounded on $B_R$.  Hence there exists an interval $\ABRACKET{a,b} \in \R$ such that $F(B_R) \subset \ABRACKET{a,b}$. Therefore $f(x)$ is continuous on an interval, i.e., a compact set, and assumes a greatest and least value for some $x_{\min}, x_{\max} \in B_R$.
	}
\end{enumerate}
\end{document}
