\documentclass[10pt,a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{hyperref}

\usepackage{multicol}
\usepackage{fancyhdr}
\usepackage[inline]{enumitem}
\usepackage{tikz}
\usepackage{tikz-cd}
\usetikzlibrary{calc}
\usetikzlibrary{shapes.geometric}
\usepackage[margin=0.5in]{geometry}
\usepackage{xcolor}

\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
    pdftitle={Tensors},
    pdfpagemode=FullScreen,
    }

%\urlstyle{same}

\newcommand{\CLASSNAME}{Math 5411 -- Mathematical Statistics I}
\newcommand{\PROFESSOR}{Nezamoddini-Kachouie}
\newcommand{\STUDENTNAME}{Paul Carmody}
\newcommand{\ASSIGNMENT}{Homework \#7 }
\newcommand{\DUEDATE}{October 18, 2024}
\newcommand{\SEMESTER}{Fall 2024}
\newcommand{\SCHEDULE}{MW 5:30 -- 6:45}
\newcommand{\ROOM}{Remote}

\pagestyle{fancy}
\fancyhf{}
\chead{ \fancyplain{}{\CLASSNAME} }
%\chead{ \fancyplain{}{\STUDENTNAME} }
\rhead{\thepage}
\input{../MyMacros}

\newcommand{\RED}[1]{\textcolor{red}{#1}}
\newcommand{\BLUE}[1]{\textcolor{blue}{#1}}

\begin{document}

\begin{center}
	\Large{\CLASSNAME -- \SEMESTER} \\
	\large{w/\PROFESSOR}
\end{center}
\begin{center}
	\STUDENTNAME \\
	\ASSIGNMENT -- \DUEDATE\\
\end{center}

\begin{enumerate}
	\item Use the definition of moment generating function (MGF) and show that:

$V[X] = d^2(M(t)) / dt^2 |_{ t = 0}  - (d(M(t)) / dt |_{ t = 0})^ 2$

\BLUE{In general
	\begin{align*}
		M^{(r)}(0) &= E\PAREN{X^r} \\
		V[X] &= E[X^2] - (E[X])^2 \\
		&= M''(0) - \PAREN{M'(0)}^2 \\
		&= d^2(M(t)) / dt^2 |_{ t = 0}  - (d(M(t)) / dt |_{ t = 0})^ 2
	\end{align*}However, there are distinct definitions of MGF for discrete and continuous random variables.
	\begin{description}
		\item Discrete Random Variables.
		\begin{align*}
			M(t) &= \sum_x e^{tx}p(x) \\
			M'(t) &= \sum_x xe^{tx}p(x) \\
			E[X] = M'(0) &= \sum_x xp(x) \\
			M''(t) &= \sum_x x^2e^{tx}p(x) \\
			E[X^2] = M''(0) &= \sum_x x^2p(x) \\
			V[X] &= \sum_x x^2p(x) - \PAREN{\sum_x xp(x)}^2 \\
		 &= E[X^2] - \PAREN{E[X]}^2 \\
		 &= M''(0)-\PAREN{M'(0)}^2
		\end{align*}
		\item Continuous Random Variables.
		\begin{align*}
			M(t) &= \int_{-\infty}^\infty e^{tx}f(x) dx \\
			M'(t) &= \int_{-\infty}^\infty xe^{tx}f(x) \\
			E[X] = M'(0) &= \int_{-\infty}^\infty xf(x) dx \\
			M''(t) &= \int_{-\infty}^\infty x^2e^{tx}f(x)dx \\
			E[X^2] = M''(0) &= \int_{-\infty}^\infty x^2f(x) dx \\
			V[X] &= \int_{-\infty}^\infty x^2f(x)dx - \PAREN{\int_{-\infty}^\infty xf(x) dx}^2 \\
		 &= E[X^2] - \PAREN{E[X]}^2 \\
		 &= M''(0)-\PAREN{M'(0)}^2
		\end{align*}
	\end{description}
}

\newpage
	\item Find the MGF for Uniform(a,b).
	
	\BLUE{
	\begin{align*}
		p(x) &= \BINDEF{\frac{1}{b-a} & a \le x \le b}{0 & \text{otherwise}} \\
		M(t) &= \int_{-\infty}^\infty e^{tx}p(x)dx \\
		&= \int_{-\infty}^a e^{tx}0dx + \int_a^b e^{tx}\frac{1}{b-a} dx+\int_b^\infty e^{tx}0 dx \\
		&= \left . \frac{1}{t(b-a)}e^{tx} \, \right |_a^b \\
		&= \frac{e^{bt}-e^{at}}{t(b-a)} \\
		M(t) &= \frac{be^{bt}-ae^{at}}{t(b-a)} 
	\end{align*}
	}
	
	\newpage
	\item Use the definitions of mean $(E[X])$ and Variance $(V[X])$ and find $E[X]$ and $V[X]$ for:
	\begin{enumerate}
		\item X~Binomial(n,p)
		
		\BLUE{\begin{align*}
			p(X=x) &= \BINDEF{p & x=1}{1-p &x=0} \\
			E[X] &= \sum_x x p(x) = 0(1-p)+1p = p \\
			E[X^2] &^= \sum_x x^2 p(x) = 0(1-p)+1p = p \\
			V[X] &= E[X^2] - (E[X])^2 = p-p^2=p(1-p)
		\end{align*}
		}

		\item X~Uniform(a,b)
		
		\BLUE{\begin{align*}
			p(X=x) &= \BINDEF{\frac{1}{b-a} & a \le x \le b}{0 & \text{otherwise}} \\
			E[X] &= \int_{-\infty}^\infty x p(x) \\
			&= \int_{-\infty}^a 0dx + \int_a^b \frac{1}{b-a}x dx+\int_b^\infty 0 dx \\
		&= \left . \frac{x^2}{2(b-a)} \, \right |_a^b \\
		&= \frac{b^2-a^2}{2(b-a)} \\
		&= \frac{b+a}{2} \\
		E[X^2] &= \int_{-\infty}^\infty x^2 p(x) \\
			&= \int_{-\infty}^a 0dx + \int_a^b \frac{1}{b-a}x^2 dx+\int_b^\infty 0 dx \\
		&= \left . \frac{x^3}{3(b-a)} \, \right |_a^b \\
		&= \frac{b^3 - a^3}{3(b-a)} \\
		&= \frac{1}{3}\PAREN{b^2+ab+a^2} \\
		V[X] &= \frac{1}{3}\PAREN{b^2+ab+a^2} - \PAREN{\frac{b+a}{2}}^2 \\
		&= \frac{1}{3}\PAREN{b^2+ab+a^2} - \frac{b^2+2ab+a^2}{4} \\
		&= \frac{1}{12}(b^2-ab+a^2)
		\end{align*}
		}
	
	\end{enumerate}

	\newpage
	\item Assume the number of Hurricanes have a Poisson distribution with average of 3 hurricanes in the hurricane season (6 months from June 1 to Nov 30). Find:

	\begin{enumerate}

		\item Probability of having no Hurricane in a hurricane season.
		
		\BLUE{\begin{align*}
			p(k) &= \frac{\lambda^k}{k!}e^{-\lambda} \AND \lambda = 3\\
			p(0) &= \frac{1}{1}e^{-3} = 0.05
		\end{align*}
		}

		\item Probability of having 1 Hurricane in a hurricane season.
		
		\BLUE{\begin{align*}
			p(1) &= \frac{1}{1} e^{-3} = 0.15
		\end{align*}
		}

		\item Probability of having 10 Hurricanes in a hurricane season.
		
		\BLUE{\begin{align*}
			p(k) &= \frac{3^{10}}{10!}e^{-3} = \frac{59049}{3628800}0.05 = 0.0008
		\end{align*}
		}

		\item Probability of having 3 Hurricanes in the first half of a hurricane season (assume the number of hurricanes are distributed evenly over 6 months of hurricane session).
		
		\BLUE{By chopping the interval in half we are effectively changing the $\lambda$ value in half or $1.5$.  Thus, our new formula would be 
		\begin{align*}
			p(k) &= \frac{1.5^k}{k!}e^{-1.5} \\
			\AND p(3) &= \frac{1.5^3}{3!}e^{-1.5} = \frac{3.375}{6}0.223 = 0.1255 
		\end{align*}
		}

\end{enumerate}
\end{enumerate}





\end{document}