\documentclass[10pt,a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{hyperref}

\usepackage{multicol}
\usepackage{fancyhdr}
\usepackage[inline]{enumitem}
\usepackage{tikz}
\usepackage{tikz-cd}
\usetikzlibrary{calc}
\usetikzlibrary{shapes.geometric}
\usepackage[margin=0.5in]{geometry}
\usepackage{xcolor}

\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
    pdftitle={Tensors},
    pdfpagemode=FullScreen,
    }

%\urlstyle{same}

\newcommand{\CLASSNAME}{Math 5411 -- Mathematical Statistics I}
\newcommand{\PROFESSOR}{Nezamoddini-Kachouie}
\newcommand{\STUDENTNAME}{Paul Carmody}
\newcommand{\ASSIGNMENT}{Homework \#4 }
\newcommand{\DUEDATE}{September 16, 2024}
\newcommand{\SEMESTER}{Fall 2024}
\newcommand{\SCHEDULE}{MW 5:30 -- 6:45}
\newcommand{\ROOM}{Remote}

\pagestyle{fancy}
\fancyhf{}
\chead{ \fancyplain{}{\CLASSNAME} }
%\chead{ \fancyplain{}{\STUDENTNAME} }
\rhead{\thepage}
\input{../MyMacros}

\newcommand{\RED}[1]{\textcolor{red}{#1}}
\newcommand{\BLUE}[1]{\textcolor{blue}{#1}}

\begin{document}

\begin{center}
	\Large{\CLASSNAME -- \SEMESTER} \\
	\large{w/\PROFESSOR}
\end{center}
\begin{center}
	\STUDENTNAME \\
	\ASSIGNMENT -- \DUEDATE\\
\end{center}

\S 42, 49, 60, and 80 from chapter 1.\\

\noindent\textbf{Exercise 42} How many ways can 11 boys on a soccer team be grouped in 4 forwards, 3 midfielders, 3 defenders, and 1 goalie?\\
\BLUE{
\begin{align*}
	\binom{11}{4,3,3,1} &= \frac{11!}{4!4!3!1!} = \frac{39916800}{24\cdot 24 \cdot 6} = \frac{39916800}{345}= 11550
\end{align*}
}

\noindent\textbf{Exercise 49} A fair coin is tossed three times
\begin{enumerate}
	\item What is the probability of two or more heads given that there were at least one head?
	\BLUE{Let $A$ be the event of "at least one head" and $B$ be the event of "two or more heads".\begin{align*}
		\Omega &= 2^3 = 8 \AND A^c = \BRACKET{ (t,t,t) }\\
		P(A) &= 7/8\\
		B^c &= \BRACKET{ (t,t,t), (h,t,t), (t,h,t), (t,t,h)}\\
		P(B) &= 1/2 \\
		B \cap A &= \BRACKET{ (h,t,t), (t,h,t), (t,t,h)}\\
		P(B\cap A) &= 3/8\\
		P(B|A) &= P(B \cap A)/P(A) = \frac{3/8}{4/8} = 3/4
	\end{align*} }
	
	\item What is the probabilty give that there was at least on tail?
	
	\BLUE{Let $C$ be the event "at least one tail".
	\begin{align*}
		P(C) &= 1/2 \\
		B\cap C &= \BRACKET{ (h,h,t), (t,h,h), (h,t,h) }\\
		P(B \cap C) &= 3/8\\
		P(B|C) &= P(B \cap C)/P(C) = \frac{3/8}{4/8} = 3/4
	\end{align*}
	}
\end{enumerate}

\noindent\textbf{Exercise 60} A factory runs three shifts.  In a given day, 1\% of the items produced by the first are defective, 2\% of the second shifts items are defective, and 5\% of the third shift's items are defective.  If the shifts all have the same productivity, what percentage of the items produced in a day are defective?  If an item is defective, what is the probability that it was produced on the third shift?\\

\BLUE{Let $N$ be the amount of items produced per shift.  The total number of defectives items $T$ will be $T = 0.01N+0.02N+0.05N = 0.08N$ will be $0.08N$ out of $3N$ items or $2.667\%$.  Let $A$ be the probability that it is a defective item and $C$ be the probability that it came from the third shift.  $P(C) = 1/3$ and $P(A) = 0.02667$.  $P(C|A) = P(C \cap A)/P(A) = 0.02667/0.05 = 0.5334$.\\
}

\noindent\textbf{Exercise 80}  If a parent has genotype A$a$, he transmits either A or $a$ to an offspring (each with a $\frac{1}{2}$ chance).  The gene he transmits to one offspring is independent of the one he transmits to another.  Consider a parent with three children and the following events:  $A = \{$children 1 and 2 have the same gene $\}$, $B=\{$ children 1 and 3 have the same gene$\}$, $C=\{$children 2 and 3 have the same gene $\}$.  Show that these events are pairwise independent but not mutually independent.
\end{document}