\documentclass[10pt,a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{hyperref}

\usepackage{multicol}
\usepackage{fancyhdr}
\usepackage[inline]{enumitem}
\usepackage{tikz}
\usepackage{tikz-cd}
\usetikzlibrary{calc}
\usetikzlibrary{shapes.geometric}
\usepackage[margin=0.5in]{geometry}
\usepackage{xcolor}
\usepackage{ esint }

\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
    pdftitle={Tensors},
    pdfpagemode=FullScreen,
    }

%\urlstyle{same}

\newcommand{\CLASSNAME}{Math 5230 -- Partial Differential Equations}
\newcommand{\STUDENTNAME}{Paul Carmody}
\newcommand{\ASSIGNMENT}{NOTES }
\newcommand{\DUEDATE}{December 17, 2025}
\newcommand{\SEMESTER}{Fall 2025}
\newcommand{\SCHEDULE}{MW 2:00 -- 3:15}
\newcommand{\ROOM}{Crawford 332}

\newcommand{\MMN}{M_{m\times n}}
\newcommand{\FF}{\mathcal{F}}

\pagestyle{fancy}
\fancyhf{}
\chead{ \fancyplain{}{\CLASSNAME} }
%\chead{ \fancyplain{}{\STUDENTNAME} }
\rhead{\thepage}
\input{../MyMacros}

\newcommand{\RED}[1]{\textcolor{red}{#1}}
\newcommand{\BLUE}[1]{\textcolor{blue}{#1}}

\begin{document}

\begin{center}
	\Large{\CLASSNAME -- \SEMESTER} \\
	\large{ w/Professor Snelson}
\end{center}
\begin{center}
	\STUDENTNAME \\
	\ASSIGNMENT -- \DUEDATE\\
\end{center} 

\begin{description}
	\item \textbf{Observations}
	
	We have the various equations: Transport, Laplace/Poisson (basic), Heat and Wave equations.  These each have general solutions, usually depicted by $\Phi$ with specific solutions based on additional information.  The PDE is the entire equation and surrounding information: initial conditions, inhomogenous results, boundary conditions.  Further, some solutions are dependent on the 'architecture/structure' of the space which they are defined (e.g., Green's Functions).  Some special cases happen requiring the isolation of "spiked points" (points that go to infinity making the PDE non-harmonic).  These include surrounding these points with a boundary exclusion zone and integrating over the smooth subspace.
	
	There are additional clever tricks that must be considered.  Integration by parts using The Divergence Theorem (building a vector space where the dot product would be useful in this).  Dirac Functions.  Green's Functions and Green's Identities.
	
	Each equation has a Maximum Value Theorem, a Mean Value Theorem, Uniqueness Theorem.

	\item \textbf{Transportation Problem and Method Characteristics}
	
	Because the Transportation Equation is
	\begin{align*}
		u_t+b\cdot Du=0
	\end{align*}resembles a linear equation we can use that as inspiration.  Use
	\begin{align*}
		z(s) = u(x+sb, t+s)
	\end{align*}we'll see that $z' = 0$ which is a constant.  Thus, when $s=-t$, intial conditions, we get $z(s) = u(x-tb, 0)$ which is our intial contioins on $u$, namely $g(x)$.  Thus, $u(x,t) = g(x-tb)$
	
	\textbf{Non-Homogenous case}: We find that $z'(s) \ne 0$ and actually is equal to $z'(s) = f(x+sb, t+s$
	\begin{align*}
		u(x,t)-g(x-bt) &= z(0)-z(-t)\\
		&= \int_{-t}^1 z'(s)\, ds \\
		&= \int_0^t f(x+(s-tb,s)\,ds
	\end{align*}thus
	\begin{align*}
		u(x,t) &= g(x-bt) +\int_0^t f(x+(s-tb,s)\,ds
	\end{align*}
	
	\item \textbf{General Laplace solution.}
\begin{remark}[\textbf{Properties of $\Phi$}].

\begin{align*}
	\Phi(x) &= \BINDEF{-\frac{1}{2\pi}\log |x| & (n=2)}{\frac{1}{n(n-2)\alpha(n)}\frac{1}{|x|^{n-2}} & (n\ge 3)} 
\end{align*}
\begin{align}
	D\Phi(y) &= \frac{-1}{n\alpha(n)}\frac{y}{|y|^n},\, (y\ne 0)\\
	\PART{\Phi}{\nu} &= \nu\cdot D\Phi(y)= \frac{-1}{n\alpha(n)\varepsilon^{n-1}},\, y \in \partial B(0,\varepsilon)  \\
	-\Delta\Phi &= \delta_0 \implies \Delta\Phi(x-y)=0,\, y \in U \AND y \ne x
\end{align}implying that there is a point source at 0 or $x$. In (2) $n\alpha(n)\varepsilon^{n-1}$ is the area of the surface of $\partial B(0,\epsilon)$

\end{remark}

	
	\item \textbf{General Poisson solution.}
	
	\begin{align*}
		&\BINDEF{-\Delta u =f, & x \in U}{u=g, & x \in \partial U}\\
		u(x,t) &= \Phi(x) - \phi^x(x) 
	\end{align*}where $\phi^x$ is a Dirichlet Delta function.
	
	\item \textbf{Green's Functions and Dirac Delta functions.}
	
	A Green's Function must satisfy the following:
	\begin{align*}
		G(x,y) &= \BINDEF{ \Delta G(x) = 0, & x \in U}{G(x,y) = \Phi(x-y) + \phi^x(x), & x,y \in \partial U}
	\end{align*}where $\phi^x$ is a Dirac Delta function determined by the geometry of $U$.  Then,
	\begin{align*}
		u(x) &= \int_U f(x)G(x,y)\, dy +\int_{\partial U} g(x)\PART{G}{\nu}(x,y) dS
	\end{align*}
	
	\item \textbf{Green's identities.}
\begin{remark}[Green's Identities].
\begin{align*}
\int_U \Delta u\; dx &= \int_{\partial U} \PART{u}{\nu} dS\\
	\int_U Dv\cdot Du\;  dx &= -\int_U u\Delta v \; dx + \int_{\partial U} \PART{v}{\nu}u\;  dS\\
	\int_U u\Delta v - v \Delta u\;  dx &= \int_{\partial U} u\PART{u}{\nu}-v\PART{u}{\nu}\;  dS
\end{align*}
\end{remark}
	
	\item \textbf{Neuman conditions.}
	
	Example from Practice Test 2:
	
	\item \textbf{Energy methods}
	
	Use the functional 
	\begin{align*}
		I[w] &= \int_U \frac{1}{2}|Dw|^2-wf\, dx \\
		w \in \mathcal{A} &= \{ w \in C^2\,|\, w=g \text{ on } \partial U\}
	\end{align*}Dirichlet's Principle then states that
	\begin{align*}
		I[u] &= \min_{w \in \mathcal{A}} I[w] \iff u \text{ is a solution}
	\end{align*}
	
	\item \textbf{Maximum Theorem}
	
	\begin{align*}
		\max_{\bar U} u &= \max_{\partial U} u
	\end{align*}
	
	\item \textbf{Mean Value Theorem}
	
	\begin{align*}
		u(x) &= \fint_U u\, dx = \fint_{\partial U} u \,dS
	\end{align*}
\end{description}

\HLINE
\begin{remark}[Integration by Parts\footnote{see page 9, 9\_17.pdf}].
\begin{align*}
	\int_U f\Delta g\, dx = -\int_U \nabla f \cdot \nabla g \, dx + \int_{\partial U} f \PART{g}{\nu}\, dS
\end{align*}
\begin{align*}
	-\Delta w &= 0 \implies w\Delta w = 0\\
	\int_U w\Delta w\; dx &= -\int_U |Dw|^2\;  dx+\int_{\partial U}wDw\cdot \nu\;  dx = 0\\
	\int_{\partial U} wDw\cdot \nu\;  dx &= \int_U |Dw|^2\;  dx + \int_U w\Delta w\;  \not dx
\end{align*}
\end{remark}

\newpage
\textbf{ChatGPT: Homework 3, question 5a: }

\begin{enumerate}[label=5(a)]
	\item Consider the initial value problem 
	\begin{align*}
		\BINDEF{u_t-\Delta u+b\cdot Du=0, & x\in \R^n, t>0}{u(x,0)=g(x) }
	\end{align*}where $b \in \R^n$ is a fixed vector. (Physically, the $b\cdot  Du$ term represents advection.)
	
	Find an explicit soluton to this probem using the  change of variables $z=x+bt$ and the solution formula for the heat equat1ion (equation (9) in Section 2.3 of Evans).
	
\end{enumerate}

\HLINE \\
Nice — this is the standard trick. The convenient choice is
$$
z := x - bt \qquad\bigl(\text{so }x=z+bt\bigr),
$$
and define
$$
v(z,t):=u(z+bt,t).
$$
Compute derivatives (keeping $z$ fixed when differentiating in $t$):
$$
v_t = u_t + b\cdot\nabla_x u,\qquad \nabla_z v=\nabla_x u,\qquad \Delta_z v=\Delta_x u.
$$
Substituting into the PDE $u_t-\Delta_x u+b\cdot\nabla_x u=0$ gives
$$
(v_t - b\cdot\nabla_x u) - \Delta_x u + b\cdot\nabla_x u = v_t-\Delta_z v =0.
$$
So $v$ satisfies the heat equation $v_t-\Delta_z v=0$ with initial data $v(z,0)=u(z,0)=g(z)$. Hence by the heat-kernel formula
$$
v(z,t)=\int_{\mathbb{R}^n}\Phi(z-y,t),g(y),dy,\qquad
\Phi(w,t)=\frac{1}{(4\pi t)^{n/2}}e^{-|w|^2/(4t)}.
$$
Returning to $u$ (recall $u(x,t)=v(x-bt,t)$) gives the explicit solution
$$
\boxed{u(x,t)=\int_{\mathbb{R}^n}\Phi(x-bt-y,t )g(y)\,dy
= \frac{1}{(4\pi t)^{n/2}} \int_{\mathbb{R}^n}
e^{-\frac{|x-y-bt|^2}{4t}}g(y)\,dy. }
$$

One can easily check this satisfies the PDE and $u(\cdot,t)\to g$ as $t\downarrow 0$. (If you prefer the change of variables $z=x+bt$ from your prompt, it is equivalent up to the sign of (b) — the above is the usual sign that removes the advection term.)

\newpage
In PDE terminology those two phrases refer to very different concepts:

---
\begin{description}

\item \textbf{ 1. **Bounded Domain**}

A *domain* is just the region in which the PDE is being considered.

* A **bounded domain** means this region is *finite in extent*.
* Example: The unit disk $ {x^2 + y^2 < 1} \subset \mathbb{R}^2 $ or the interval $ (0,1) \subset \mathbb{R} $.

This term is purely **geometric** — it says nothing about time or the type of PDE.

---

\item  \textbf{2. **Parabolic Boundary**}

This term is specific to **evolution PDEs**, especially **parabolic equations** like the heat equation.

Suppose you solve a PDE on a space-time cylinder:

$$
\Omega \times (0,T),\quad \text{where }\Omega\subset\mathbb{R}^n\text{ is the spatial domain}.
$$

The **parabolic boundary** $ \partial_p(\Omega \times (0,T)) $ consists of:

* The **initial time boundary**: $ \Omega \times {0} $
* The **lateral spatial boundary over time**: $ \partial\Omega \times (0,T) $

So for the heat equation, you don’t impose boundary conditions on the *final time* $t=T$, because information flows *forward* in time.

---

\item  \textbf{Summary Table}

\begin{tabular}{|l|l|l|l|}
\hline
Term  & Applies To  & Meaning  & Example \\
\hline
**Bounded domain**  & Any PDE  & Spatial region is finite & $ \Omega = (0,1) \subset \mathbb{R} $ \\
**Parabolic boundary** & Parabolic PDEs (e.g. heat equation) & Where boundary/initial data & $ \Omega\times{0} \cup \partial\Omega \times (0,T) $ \\
& &must be given in space-time & \\
\hline
\end{tabular}

\end{description}

Let me know if you'd like a sketch or diagram — this is much easier to see visually!


\end{document}