\documentclass[10pt,a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{hyperref}

\usepackage{multicol}
\usepackage{fancyhdr}
\usepackage[inline]{enumitem}
\usepackage{tikz}
\usepackage{tikz-cd}
\usetikzlibrary{calc}
\usetikzlibrary{shapes.geometric}
\usepackage[margin=0.5in]{geometry}
\usepackage{xcolor}

\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
    pdftitle={Tensors},
    pdfpagemode=FullScreen,
    }

%\urlstyle{same}

\newcommand{\CLASSNAME}{Math 5111 -- Real Analysis II}
\newcommand{\STUDENTNAME}{Paul Carmody}
\newcommand{\ASSIGNMENT}{Semester Notes }
\newcommand{\DUEDATE}{May 2025}
\newcommand{\SEMESTER}{Sprint 2025}
\newcommand{\SCHEDULE}{MW 9:30 - 10:45}
\newcommand{\ROOM}{Remote}

\newcommand{\MMN}{M_{m\times n}}
\newcommand{\FF}{\mathcal{F}}

\pagestyle{fancy}
\fancyhf{}
\chead{ \fancyplain{}{\CLASSNAME} }
%\chead{ \fancyplain{}{\STUDENTNAME} }
\rhead{\thepage}
\input{../MyMacros}

\newcommand{\RED}[1]{\textcolor{red}{#1}}
\newcommand{\BLUE}[1]{\textcolor{blue}{#1}}
\newcommand{\ts}{\textsuperscript}

\begin{document}

\begin{center}
	\Large{\CLASSNAME -- \SEMESTER} \\
	\large{ w/Professor Liu}
\end{center}
\begin{center}
	\STUDENTNAME \\
	\ASSIGNMENT -- \DUEDATE\\
\end{center} 

\noindent January 18\ts{th}, 2025\\

\noindent \textbf{Continuity} is a property unlike most that we encounter in understanding functions.  Typically, we start with the domain and apply it to the function and get a result in the range.  A function is continuous, however, when the area around the range has the same properties as the area around the domain.
\begin{enumerate}
    \item  \textbf{continuous functions} defined using epsilon/delta.  Which becomes increasingly difficult when our domain and range go beyond the Real numbers (i.e., multiple dimensions, complex numbers, sets that aren’t compact).
    \item  \textbf{Topological definition of continuous functions} which is that an open set in the range comes from an open set in the domain.  This skips over measure by simply redefining “open sets” in more abstract terms.  A metric topology has a way of measuring distance but it isn’t necessary for this definition.
    \item  Analogously speaking, \textbf{a function is measurable} (a.k.a., continuous) when a measurable (a.ka., open) set in the range comes from a measurable (a.k.a., open) set in the domain.  The sets are defined by the Lebesgue Outer Measure (basically, ‘not a point’, which is much like but not the same as an open set in topology).
    \item Definitions of "superset" topology and $\sigma$-algebra.  These two concepts are so similar that they might be the same thing.
    \begin{enumerate}
    		\item A \textbf{topology} is defined as the collection where at least the set and the empty set as members, as well as subsets where all intersections (countable) and all unions (even uncountable) are also members.  "Open sets" are simply the members of the topology.
    		\item A $\sigma$-algebra is defined as having the set itself, all compliments of its members (which are subsets), and countable unions of its members.\footnote{$\sigma$ signifies inifinite unions and $\delta$ (Not shown) would signify infinite intersectios} The set and $\sigma$-algebra combine to make a \textbf{Measurable Space} and its members are called \textbf{Measurable Sets} (analogous to "open sets"). If $f$ maps from a measurable space to a topological sapce and each open in the range comes from measurable in the domain (i.e., $f^{-1}(V)$ maps open set $V$ from a measurable set through $f$) then $f$ is said to be \textbf{Measurable Function}.
    		\item These two appear to me to be logically equivalent.
	\end{enumerate}      
    \item  Further, a \textbf{Borel Set} is an element of the smallest possible $\sigma$-algebra generated by the set.  It is, consequently, measurable (in the Lebesgue sense, i.e., ‘not a point’).  And a function is \textbf{Borel Measurable} when Borel Sets are mapped from Borel Sets.

\end{enumerate}

\noindent Rules for Composing different types of these functions.

\begin{tabular}{|l|c|c|c|c|}
\hline 
	& Nomen.\footnote{name of the set} & $f:X \to Y$ & $g:Y \to Z$ & $g\circ f(x) =g(f(x)) = z$ \\
	\hline
	standard & open inteval &continuous & continuous & continuous \\
	\hline
	topology\footnote{analogous to standard} & open &continuous & continuous & continuous \\
	\hline
	measurable & measurable & measurable & continuous & measurable \\
	\hline
	measurable & measurable & measurable & measurable & measurable \\
	\hline 
	Borel Measurable & Borel & Borel measurable  & continuous & ??? \\
	\hline 
	Borel Measurable & Borel & continuous & Borel measurable & ??? \\
	\hline 
	Borel Measurable & Borel & measurable & Borel mapping & measurable \\
	\hline 
	Borel Measurable & Borel & Boreal measurable & Borel mapping & Borel measurable \\
\hline
\end{tabular}
\end{document}
