\documentclass[12pt,a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{hyperref}

\usepackage{multicol}
\usepackage{fancyhdr}
\usepackage[inline]{enumitem}
\usepackage{tikz}
\usepackage{tikz-cd}
\usetikzlibrary{calc}
\usetikzlibrary{shapes.geometric}
\usetikzlibrary{positioning}
\usepackage[margin=0.5in]{geometry}
\usepackage{xcolor}

\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
    pdftitle={Tensors},
    pdfpagemode=FullScreen,
    }

%\urlstyle{same}

\newcommand{\CLASSNAME}{Math 5050 -- Special Topics: Manifolds}
\newcommand{\STUDENTNAME}{Paul Carmody}
\newcommand{\ASSIGNMENT}{Section 8: The Tangent Space }
\newcommand{\DUEDATE}{May 30, 2025}
\newcommand{\PROFESSOR}{Professor Berchenko-Kogan}
\newcommand{\SEMESTER}{Fall 2025}
\newcommand{\SCHEDULE}{TBD}
\newcommand{\ROOM}{Remote}

\newcommand{\MMN}{M_{m\times n}}
\newcommand{\FF}{\mathcal{F}}

\pagestyle{fancy}
\fancyhf{}
\chead{ \fancyplain{}{\CLASSNAME} }
%\chead{ \fancyplain{}{\STUDENTNAME} }
\rhead{\thepage}
\input{../MyMacros}

\newcommand{\RED}[1]{\textcolor{red}{#1}}
\newcommand{\BLUE}[1]{\textcolor{blue}{#1}}

\begin{document}

\begin{center}
	\Large{\CLASSNAME -- \SEMESTER} \\
	\large{ w/\PROFESSOR}
\end{center}
\begin{center}
	\STUDENTNAME \\
	\ASSIGNMENT -- \DUEDATE\\
\end{center} 


\noindent \textbf{\\\large{Problems}}

\begin{enumerate}[label=9.\arabic*.]

\item \textbf{Regular values}

Define $f:\R^2\to \R$ by
\begin{align*}
	f(x,y) = x^3-6xy+y^2.
\end{align*}Fine all values $c \in \R$ for whcih the level $f^{-1}(c)$ is a regular submanifold of $\R^2$.

\item \textbf{Solution set of one equation.}

Let $x,y,z,w$ be the standard coordinates on $\R^4$.  Is teh solution set of $x^5+y^5+z^5+w^5=1$ in $\R^4$ a smooth manifold?  Explain why or why not.  (Assume that the subset is given the subspace topology).

\item \textbf{Solution set of two equations.}

Is teh solution set of thesysemt of equations 
\begin{align*}
	x^3+y^3+z^3=1,\, z=xy
\end{align*}in $\R^3$ a smooth manifold?  Prove your answer.

\item \textbf{Regular submanifolds}

Suppose that a subset $S$ of $\R^2$ hat eh property that locally on $S$ one of the coordinates is $C^\infty$ function of the other coordinate.  Show that $S$ is qa regular submanifold of $\R^2$. (Note that the unit circle defined by $x^2+y^2=1$ has this property. AT every point of the circle, there is a neighborhood in whic h$y$ is a $C^\infty$ function of $x$ or $x$ is a $C^\infty$ function of $y$.)

\item \textbf{Graph of a smooth function}

Show that the graph $\Gamma(f)$ of a smooth function $f: \R^2\to \R$.
\begin{align*}
	\Gamma(f) &= \BRACKET{\PAREN{x,y,f(x,y)} \in \R^3}
\end{align*}is a regular submanifold of $\R^3$.

\item \textbf{Euler's formula}
A polynomial $F(x_0,\dots,x_n) \in \R[x_0,\dots,x_n]$ is \textit{homogenous of degree $k$} if it is a linear combination of monomials $x_9^{i_o}\cdots x_n^{i_n}$ of degree $\SUM{j=0}{n}i+j = k$.  Let $F(x_0,\dots,x_n)$ be a homogenous polynomial of degree $k$.  Clearly, for any $t \in \R$,
\begin{align*}
	F(tx_o,\dots,tx_n) &= t^kF(x_0,\dots, x_n).
\end{align*}Show that
\begin{align*}
	\sum+{i=0}^n x_i\PART{F}{x_i} = kF.
\end{align*}

\item \textit{Smooth projectve hypersurface}

On the projective space $\R P^n$ a hmogenous polynomial $F(x_0,\dots,x_n)$ of degree $k$ is not a function , since its value at a point $[a_0,\dots,a_n]$ is not unique.  However, the zer set in $\R P^n$ of a homogenous polynomial $F(x_0,\dots,x_n)$ is well defined, since $F(a_0,\dots,a_n)=0$ if and only if 
\begin{align*}
	F(ta_0,\dots,ta_n) &= t^kF(a_0,\dots,a_n)=0,\, \forall t \in R^\times: \R-\{0\}
\end{align*}The zero set of finitely many homogenous polynomials in $\R P^q$ is called a \textit{real projective variety}.  A projectvie variety defined bya single homogeneous polynomial of degree $k$ is called a \textit{hypersurface} of degree $k$.  Show that the hypersurface $Z(F)$ defined by $F(x_0,x_1,x_2)=0$ is smoot if $\partial F/\partial x_0, \partial F/\partial x_1$ and $\partial F/\partial x_2$ are simultaneously zero on $Z(F)$.  (\textit{Hint:} The standard coordinates on $U_0$ which is homeomorpic to $\R^2$, are $x=x_1/x_0, y=x_2/x_0$ (see Subsection 7.7).  In $U_0, F(x_0,x_1,x_2) = x_0^tF(1,x_1/x_0, x_2/x_0)= x_0^kF(1,x,y)$.  Defein $f(x,y)= F(1,x,y)$.  Then $f$ and $F$ ahve the same zero set in $U_0$.)

\item \textbf{Product of regular submanifolds}

If $S_1$ is a regular submanifold of the manifold $M_i$ for $i=1,2$, prove that $S_1 \times S_2$ is a regular submanifold of $M_1 \times M_2$.

\end{enumerate}
\end{document}
