\documentclass[10pt,a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{hyperref}

\usepackage{multicol}
\usepackage{fancyhdr}
\usepackage[inline]{enumitem}
\usepackage{tikz}
\usepackage{tikz-cd}
\usetikzlibrary{calc}
\usetikzlibrary{shapes.geometric}
\usetikzlibrary{positioning}
\usepackage[margin=0.5in]{geometry}
\usepackage{xcolor}

\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
    pdftitle={Tensors},
    pdfpagemode=FullScreen,
    }

%\urlstyle{same}

\newcommand{\CLASSNAME}{Math 5050 -- Special Topics: Manifolds}
\newcommand{\STUDENTNAME}{Paul Carmody}
\newcommand{\ASSIGNMENT}{Section 7: Quotients }
\newcommand{\DUEDATE}{May 17, 2025}
\newcommand{\PROFESSOR}{Professor Berchenko-Kogan}
\newcommand{\SEMESTER}{Fall 2025}
\newcommand{\SCHEDULE}{TBD}
\newcommand{\ROOM}{Remote}

\newcommand{\MMN}{M_{m\times n}}
\newcommand{\FF}{\mathcal{F}}

\pagestyle{fancy}
\fancyhf{}
\chead{ \fancyplain{}{\CLASSNAME} }
%\chead{ \fancyplain{}{\STUDENTNAME} }
\rhead{\thepage}
\input{../MyMacros}

\newcommand{\RED}[1]{\textcolor{red}{#1}}
\newcommand{\BLUE}[1]{\textcolor{blue}{#1}}

\begin{document}

\begin{center}
	\Large{\CLASSNAME -- \SEMESTER} \\
	\large{ w/\PROFESSOR}
\end{center}
\begin{center}
	\STUDENTNAME \\
	\ASSIGNMENT -- \DUEDATE\\
\end{center} 

\noindent\textbf{Pg. 77: Exercise 7.11 (Real projective space as a quotient of a sphere).*}  For $x=(x^1,\dots,x^n)\in \R^n$, let $||x|| = \sqrt{\sum_i(x^i)^2}$ be the modulus of $x$.  Prove that the map $f: \R^{n+1} - \{0\} \to S^n$ given by 
\begin{align*}
	f(x) = \frac{x}{||x||}
\end{align*}induces a homeomorphism $\bar{f}: \R P^n\to S^n/\sim$.  Where
\begin{align*}
	x \sim y &\iff x=\pm y, \, x,y \in S^n
\end{align*} (\textit{Hint: } Find an inverse map 
\begin{align*}
	\bar{g}: S^n/\sim \,\to \R P^n
\end{align*}and show that both $\bar{f}$ and $\bar{g}$ are continuous.)\\

\BLUE{Given the relation above $x \sim y \iff x=\pm y,\, x,y \in S^n$.  Define
	\begin{align*}
		\bar{g}([x]) &= [x] 
	\end{align*}Note that on the left $[x] \in S^n/\sim$ and $[x] \in \R P^n$.  For clarity,\footnote{$\R P^n \equiv \R^{n+1}/r(x,y)$ where $r(x,y)$ is the relation that is true when $x,y,p$ are colinear.} 
	\begin{align*}
		[a] \in S^n/\sim &\implies [a] = \{ a, -a\} \WHERE a \in S^n\\
		[b] \in \R P^n &\implies [b] = \{x \in \R^{n+1}|\, x=\alpha b,\, \forall \alpha \in \R\} \WHERE b \in \R^{n+1} 
	\end{align*}Notice that
	\begin{align*}
		\bar{f}([b]) &= [f(b)] = \SQBRACKET{\frac{b}{||b||}} \in S^n/\sim \\
		\bar{g}\PAREN{\SQBRACKET{\frac{b}{||b||}}} &= [b] \\
		\therefore \bar{g}\circ\bar{f} = \mathbb{I}\\
		\AND \bar{g}([a]) &= [a] \in \R P^n \\
		\bar{f}\PAREN{\bar{g}([a])} &= \SQBRACKET{\frac{[a]}{|| [a]||}} = [f(a)] = [a] \in S^n/\sim\\
		\therefore \bar{f}\circ\bar{g} = \mathbb{I}
	\end{align*} $\bar{f}$ is continuous because $f$ is continuous.  $\bar{g}$ is continuous because it is a mapping of one identity to another.  Therefore, $\bar{f}$ is a homeomorphism.
	}
	
\HLINE
\begin{tikzpicture}
	\node (S) {$S$};
	\node (Y)  [right=of S] {$Y$};
	\node (Q)  [below= of S] {$S/\sim$};
	\draw (S) -> node[above] {$f$} (Y) 
		(S) ->  node[left] {$\pi$} (Q)
		(Q) -> node[right] {$\hat{f}$} (Y);
\end{tikzpicture}\\
\textbf{Problems}

\begin{enumerate}[label=7.\arabic*.]
	\item \textbf{Image of the inverse image of a map}
	
	Let $f:X\to Y$ be a map of sets, and let $B \subset Y$.  Prove that $f(f^{-1}(B)) = B \cap f(X)$.  Therefore, if $f$ is surjective, then $f(f^{-1}(B)) = B$.
	
	\BLUE{\textbf{$\subseteq$:  }Let $b \in B$ and $a \in X$ such that $f(a) = b$.  Then, $a \in f^{-1}(b)$, thus $a$ is an arbitrary point in $f^{-1}(B)$.  We know that $f(a) \in f(X)$ and $f(a) \in B$, therefore $f(a) \in B \cap f(X)$ and $f(f^{-1}(B)) \subseteq B \cap f(X)$.\\ \\
	\textbf{$\supseteq$:  }Let $b \in B \cap  f(X)$.  Since $b \in f(X)$, there exists $a \in X$ such that $f(a) = b$ and since $b \in B$ then $b = f(a) \in f(f^{-1}(b)) \subseteq f^{-1}(B)$.  Therefore $b \in f(f^{-1}(B))$\\
	}
	
	\item \textbf{Real projective plane}
	
	Let $H^2$ be the closed upper hemisphere in the unit sphere $S^2$, and let $i: H^2 \to S^2$ be the inclusion map.  In the notation of example 7.13, prove that the induced map $f: H^2/\sim \,\to S^2/\sim$ is a homeomorphism. (\textit{Hint:} Imitate Propostion 7.3.)
\newcommand{\HOMEO}{\overset{\sim}\to}	
	\BLUE{Let $H^2$ be the upper hemisphere and $S^2$ be the unit sphere
	\begin{align*}
		H^2 &= \{(x,y,z) \in \R^3\,|\, x^2+y^2+z^2=1, z \ge 0\} \\
		S^2 &= \{(x,y,z) \in \R^3\,|\, x^2+y^2+z^2=1\}
	\end{align*}These two are homeomorphic to each via 
	\begin{align*}
		\varphi : S^2 &\mapsto H^2 \\
		\varphi(x,y,z) &= (x,y, |z|)
	\end{align*} and its inverse
	\begin{align*}
		\psi : H^2 &\mapsto S^2 \\
		\psi(x,y,z) &= (x,y,z)
	\end{align*}Define the relations
	\begin{align*}
		(x,y,z) \sim (-x,-y,-z) &\to x^2+y^2=z^2, \, \forall (x,y,z) \in S^2 \\
		(x,y,z) \sim (x,y, z) &\to \sqrt{x^2+y^2}=z, \, \forall (x,y,z) \in H^2
	\end{align*}Then we induce $f:H^2/\sim\, \to S^2/\sim$ as
	\begin{align*}
		f([(x,y,z)]) &= [(x,y,z)]
	\end{align*}
	}
	
	\item \textbf{Closedness of the diagonal of a Hausdorff Space}
	
	Deduce Theorem 7.7 from Corollary 7.8 (\textit{Hint:} To prove that if $S/\sim$ is Hausdorff, then the graph $R$ of $\sim$ is closed in $S \times S$, use the continuity of the projection map $\pi:S \to S/\sim$.  To prove the reverse implication, use the openness of $\pi$.)
	
	\BLUE{Assuming Corollary 7.8, that a Topological Spacee is Hausdorf if and only if the diagonal $\Delta$ in $S \times S$ closed.  Thus, $R = S \times S - \Delta$ must be open.  That means there exists and $U \times V \subset S \times S - \Delta$ where $U, V \in S$ such that $U \cap V = \emptyset$.  Since $\pi$ is continuous, let $A=\pi^{-1}(U)$ and $B=\pi^{-1}(V)$.  Clearly $A \cap B = \emptyset$.  Thus, $x\in A$ and $y \in B$ means that $x\ne y$. Let $\sim \equiv =$. Then, there exists $[x], [y] \in S/\sim$ where $[x] \cap [y] = \emptyset$
	}
	
	\item \textbf{Quotient of a sphere with antipodal points indentified}
	
	Let $S^n$ be the unit sphere centered at the origin $\R^{n+1}$.  Define an equivalence relation $\sim$ on $S^n$ by indentifying antipodal points:
	\begin{align*}
		x \sim y \iff x=\pm y, \, x,y, \in S^n
	\end{align*}\begin{enumerate}[label=(\alph*)]
		\item Show that $\sim$ is an open equivalence relation.

		\BLUE{Given an open set $U$ and $x \in U$.  Then there exists $\epsilon > 0$ such that the $n$ dimensional ball $B(x,\epsilon)=\{y: |y -x| < \epsilon\} \subset U$.  Let $z \in B(x,\epsilon)$ and $z \sim z'$.  Then, $z = \pm z'$ and $|z -x |< \epsilon$ implies that $|\pm z'-x|< \epsilon$ or $z' \in B(x, \epsilon)$.  $z,z'$ are arbitrary points in $B(x,\epsilon)$ thus the open ball is mapped to itself and is open.  Therefore $\sim$ is open.
		}
		
		\item Apply Theorem 7.7 and Corallary 7.8 to prove that the quotient space $S^n/\sim$ is Hausdorff, without making use of the homeomorphism $\R P^n \cong S^n/\sim$.
		
		\BLUE{From (a), $\sim$ is an open relation.  WTS that the graph $R$ of $\sim$ in $S \times S$ is closed, thus, from 7.8, $S^n/\sim$ is Hausdorff.  We can do this by showing that the diagonal of $R$, $\Delta = \{(x,x) \in S^n \times S^n\}$, is closed.\\
		Let $P = \Delta^c = S^n \times S^n - \Delta$.  Let $(U,V) \subseteq P$ for open sets $U,V \subseteq S$.  If $U \cap V = \emptyset$ then $U \sim V = \emptyset$ is open. Otherwise, $U \cap V$ is open.  Define $Q=\{ (x,y) \in U \cap V\,|\,x\sim y\}$.  Since $\sim$ is an open relations and $U \cap V$ is an open set, $Q$ must be open. $Q \subseteq P$.  Now, let $N = \{(x,y) \in P - Q\}$.  Then $x \ne y$ and $x \nsim y$.
		}
		
	\end{enumerate}
	
	\item \textbf{Orbit space of a continuous group action.}
	
	Suppose a right action of a topological group $G$ on a topological space $S$ is continuous, this simply means that the map $S \times G \to S$ describing the action is continuous.  Define two points $x,y$ of $S$ to be equivalent if they are in the same orbit; i.e., there is an element $g \in G$ such that $y=xg$.  Let $S/G$ be the quotient space; it is called the \textit{orbit space} of the action.  Prove that the projection map $\pi: S \to S/G$ is an open map.  (This problem generalizes Proposition 7.14, in which $G=R^\times = \R-\{0\}$ and $S = \R^{n+1}-\{0\}$.  Because $\R^\times$ is commutative, a left $\R^\times$-action becomes a right $\R^\times$-action if scalar multiplication is written on the right.)
	
	\BLUE{$\dagger$Want to show that given an open set $U \in S$, then $\pi^{-1}(\pi(U)) \in S$ is open  given that there exists $f: S\times G \to S$ which is continuous.
	\begin{align*}
		\pi^{-1}(\pi(U)) &= \{ x \in S\,|\, \pi(x) \in \pi(U)\} \\ 
		&= \{ x \in S\,|\, [x] \cap U \ne \emptyset \}\\
		&= \bigcup_{g \in G} Ug
	\end{align*}Thus, this mapping $\pi^{-1} \circ \pi$ on an open set is a set of elements in $S$ based on $xg$ for all $x \in U$ and $g \in G$.  Now $f(x,g): S\times G \to S$ is continuous and $f(x,g) =xg$.  Thus $f^{-1}(U)$ is open for all $g \in G$.  Thus each $Ug$ is open in $S$.  $(\pi^{-1}\circ \pi)(U)$ is the union of open sets and is therefore open.
	}
	
	\item \textbf{Quotient of $\R$ by $2\pi\Z$.}
	
	Let the additive group $2\Z$ act on $\R$ on the right by $x\cdot 2\pi n=x+2\pi n$, where $n$ is an integer.  Show that the orbit space $\R/2\pi\Z$ is a smooth manifold.
	
	\BLUE{Primary Option: Let $\R = S$ and $G = 2\pi \Z$ and $[x] = \{x\in \R: x + 2 \pi n, \forall n \in \Z\}$ .  \\
	Second Option: Must show that it is 2\SND countable, Hausdorff and there exists an atlas covering the set.
	\begin{description}
	\item 2\SND countable.\\	
	Yes. The basis for $\R$ is a basis for $\R/2\pi\Z$.
	\item Hausdorff.\\
	Yes.  Every open set in $\R/2\pi \Z$ is open in $\R$ and every point as well.  Thus, Hausdorff.
	\item An Atlas Exists.\\
	Yes.  Let $U_i = ( \pi i, \pi i + 2 \pi )$ and $\phi_i(x) = x - 2 \pi i$. And $\bigcup_{i\in\Z} U_i = \R$ and $\phi_i \in \C^\infty$
	\end{description}
	}
	
	\item \textbf{The circle as a quotient space}
	
	\begin{enumerate}[label=(\alph*)]
	
		\item Let $\{(U_\alpha, \phi_\alpha)_{\alpha=1}^2\}$ be the atlas of circles $S^1$ in Example 5.7, and let $\bar{\phi}_\alpha$ be the map $\phi_\alpha$ followed by the projection $\R \to \R/2\pi\Z$.  On $U_1\cap U_2 = A \coprod B$, since $\phi_1$ and $\phi_2$ differ by an integer multiple of $2\pi$, $\bar{\phi}_1 = \bar{\phi}_2$.  Therefore, $\bar{\phi}_1$ and $\bar{\phi}_2$ piece together to give a well-defined map $\bar{\phi}: S^1\to \R/2\pi\Z$.  Prove that $\bar{\phi}$ is $C^\infty$.
		
		\BLUE{From Example 5.7
		\begin{align*}
			U_1 &= \{e^{it} \in \C\,|\, -\pi < t < \pi \} \AND \phi_1(e^{it}) = t, \, -\pi < t < \pi\\
			U_2 &= \{e^{it} \in \C\,|\, 0 < t < 2\pi\} \AND \phi_2(e^{it}) = t,\, 0 < t < 2\pi \\
			A &= \{e^{it}\,|\, -\pi < t < 0\}\\
			B &= \{e^{it}\,|\, 0 < t < \pi\} 	\\
			\bar{\phi}_1(e^{i(t+2\pi n)}) &= [ \phi_1(e^{it}e^{2\pi n})] = [ \phi_1(e^{it})]=[t],\, -\pi < t < \pi\\
			\bar{\phi}_2(e^{it}) &= [ \phi_2(e^{it})] = [t],\, 0 < t < 2\pi \\
			\bar{\phi}(e^{it}) &= [t] \in \R/2\pi\Z, \, \forall t \in \R
		\end{align*}
		}
		
		\item The complex exponential $\R \to S^1, t\mapsto e^{it}$, is constant on each orbit of the action of $2\pi\Z$ on $\R$.  Therefore, there is an induced map $F: \R/2\pi\Z \to S^1, F([t])=e^{it}$.  Prove that $F$ is $C^\infty$.
		
		\BLUE{\begin{tikzpicture}
	\node (S) {$\R$};
	\node (Y)  [right=of S] {$S^1$};
	\node (Q)  [below= of S] {$\R/2\pi \Z$};
	\draw [->] (S) -> node[above] {$f$} (Y);
	\draw [->] (S) ->  node[left] {$\pi$} (Q);
	\draw [->] (Q) -> node[right] {$F$} (Y);
\end{tikzpicture}\\
$f: \R \to S^1, t \mapsto e^{it}$ is smooth, and $\pi: \R \to \R/2\pi \Z, t \mapsto [t]$ is smooth, from the diagram above $f = F \circ \pi$.  Therefore $F$ is smooth.
		}
		
		\item Prove that $F: \R/2\pi\Z \to S^1$ is a diffeomorphism.
	
	\end{enumerate}
	
	\item \textbf{The Grassmanian $G(k,n)$}
	
	\item \textbf{Compactness of real projective space}
	
	Show that the real projective space $\R P^n$ is compact. (\textit{Hint:} Use Exercise 7.11.)
	

\end{enumerate}
\end{document}
